---
title: "Homework 2"
output: github_document
---
```{r setup}
library(tidyverse)
library(readxl)
```

## Problem 1

Read and clean the Mr. Trash Wheel sheet. 

```{r}
trashwheel_df = 
  read_xlsx("./data/trash_wheel_collection.xlsx",
    sheet = "Mr. Trash Wheel",   
    range = cell_cols("A:N")) %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster) %>% 
  mutate(
    sports_balls = round(sports_balls),
    sports_balls = as.integer(sports_balls)
  )
```

Read precipitation data for 2018 and 2017.

```{r}
precip_2018 = 
  read_excel(
    "./data/trash_wheel_collection.xlsx", 
    sheet = "2018 Precipitation", 
    skip = 1, 
  ) %>% 
  janitor::clean_names() %>% 
  drop_na(month) %>% 
  mutate(year = 2018) %>%
  relocate(year)

precip_2017 = 
  read_excel(
    "./data/trash_wheel_collection.xlsx", 
    sheet = "2017 Precipitation", 
    skip = 1, 
  ) %>% 
  janitor::clean_names() %>% 
  drop_na(month) %>% 
  mutate(year = 2017) %>%
  relocate(year)
```


Now combine annual precipitation. 

```{r}
month_df = 
  tibble(
    month = 1:12, 
    month_name = month.name
  )

precip_df = 
  bind_rows(precip_2018, precip_2017)

left_join(precip_df, month_df, by = "month")

```

This dataset contains information from the Mr. Trash Wheel trash collector in Baltimore, Maryland. As trash enters the inner harbor, the trash wheel collects that trash and stores it in a dumpster. The dataset contains information on year, month, and trash collected, including some specific kinds of trash. There are a total of `r ncol(trashwheel_df)` columns and `r nrow(trashwheel_df)` rows in our final dataset. Additional data sheets include month and precipitation data for each year between 2014-2019. 

The total precipitation in 2018 was 

The median number of sports balls in a dumpster in 2017 was 






