---
title: "Homework 2"
output: github_document
---
```{r setup}
library(tidyverse)
library(readxl)
```

## Problem 1: Mr. Trash Wheel

Read and clean the Mr. Trash Wheel sheet. 

```{r}
trashwheel_df = 
  read_xlsx("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "Mr. Trash Wheel",   
    range = cell_cols("A:N")) %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster) %>% 
  mutate(
    sports_balls = round(sports_balls),
    sports_balls = as.integer(sports_balls)
  )
```

Read and clean precipitation data for 2018 and 2017.

```{r}
precip_2018 = 
  read_xlsx(
    "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", 
    sheet = "2018 Precipitation", 
    skip = 1, 
  ) %>% 
  janitor::clean_names() %>% 
  drop_na(month) %>% 
  mutate(year = 2018,
         total = as.numeric(total)) %>%
  relocate(year)

precip_2017 = 
  read_xlsx(
    "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", 
    sheet = "2017 Precipitation", 
    skip = 1, 
  ) %>% 
  janitor::clean_names() %>% 
  drop_na(month) %>% 
  mutate(year = 2017) %>%
  relocate(year)
```


Now combine annual precipitation and change month number to name by creating separate month tibble.  

```{r}
month_df = 
  tibble(
    month = 1:12, 
    month_name = month.name
  )

precip_df = 
  bind_rows(precip_2018, precip_2017)

left_join(precip_df, month_df, by = "month")

```

This dataset contains information from the Mr. Trash Wheel trash collector in Baltimore, Maryland. As trash enters the inner harbor, the trash wheel collects that trash and stores it in a dumpster. The dataset contains information on year, month, and trash collected, including some specific kinds of trash. There are a total of `r ncol(trashwheel_df)` columns and `r nrow(trashwheel_df)` rows in our final dataset. Additional data sheets include month and precipitation data for each year between 2014-2019. 

The total precipitation in 2018 was `r sum(pull(precip_2018, total))`inches. 

The median number of sports balls in a dumpster in 2017 was `r median(pull(filter(trashwheel_df, year == 2017), sports_balls))` balls.




## Problem 2: NYC Transit

*Part 1: Read and clean the NYC Transit data.*

```{r}
nyc_transit_df = 
  read_csv("./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>% 
  janitor::clean_names() %>% 
  select(line:entry, vending, ada) %>%
  mutate(entry = recode(entry, "YES" = "TRUE", "NO" = "FALSE"))

view(nyc_transit_df)
```

The NYC Transit dataset contains a variety of variables pertaining to the NYC transit system. After reading in the csv file, I cleaned the data by first using the clean_names function from the janitor package to make all variable names in the same format with lowercase letters and snake case. Then, I used the select command to retain only the variables of interest and omit all other extraneous variables. The following variables are those of interest: line, station, name, station latitude and longitude, routes served, entry, entrance type, vending, and ADA compliance. Finally, I converted the "entry" variable from the original character vector into a logical vector by using the recode function embedded in the mutate function. The resulting dataset contains `r nrow(nyc_transit_df)` rows and `r ncol(nyc_transit_df)` columns. The data is relatively tidy since each variable and entry has its own column and row, respectively. 


*Part 2: Details about the data.*



